\PassOptionsToPackage{unicode,pdfusetitle}{hyperref}
\PassOptionsToPackage{hyphens}{xurl}
\PassOptionsToPackage{usenames,dvipsnames}{xcolor}

\documentclass[english,a4paper,14pt]{letter}

\usepackage{babel}

% Font settings
\usepackage[T1]{fontenc}
\usepackage{libertine}
\usepackage{inconsolata}

\usepackage{textcomp}
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath}

\usepackage{xcolor}

% tables and lists
\usepackage{enumitem}
\usepackage{booktabs}

\usepackage{graphicx}

\usepackage{xurl}

\usepackage{fancyhdr}
\fancyhead{}
\fancyhead[L]{Larsson, Bogdan, Grzesiak, Massias, Wallin}
\fancyhead[R]{Cover Letter}

\pagestyle{fancy}
\thispagestyle{plain}

\usepackage{hyperref}
\hypersetup{
  pdftitle   = {Cover Letter},
  pdfauthor  = {Johan Larsson, Malgorzata Bogdan, Krystyna Grzesiak, Mathurin Massias, Jonas Wallin},
  colorlinks = true,
  linkcolor  = MidnightBlue,
  citecolor  = MidnightBlue,
  urlcolor   = MidnightBlue,
  breaklinks = true
}

\begin{document}

\begin{letter}{Journal of Statistical Software}
  \opening{Dear Editor,}

  We are thankful for the feedback on our previous submission and apologize for the
  issues encountered in the submission. Below, we provide a point-by-point response to the
  comments raised and outline the changes made to address them.

  We would, however, like to take this opportunity to clear up what we think might
  be a misunderstanding regarding the software and its scope. SLOPE solves
  a type of regularized regression problem so it is not related to \emph{slopes}
  in the geometric sense. Furthermore, SLOPE is distinct from standard \(\ell_1\)-penalized
  regression (the lasso), so we do not see the relevance of a full background
  on packages that implement the lasso. We do, however, realize that we
  neglected to discuss existing SLOPE implementations, even if to explain
  that there are in fact no other specific SLOPE software packages, and have
  added such a discussion in the introduction.

  We hope that you will find our revised submission suitable for publication in
  the Journal of Statistical Software.

  Here are the detailed comments and our answers:

  \begin{quote}
    Journal of Statistical Software requires that the software
    usage is illustrated on simple problems run step-by-step in the
    manuscript, showing command lines and their results. This is currently
    done for the R package but we would find beneficial to have a similar
    code for the other two packages.
  \end{quote}

  \begin{quote}
    The introduction must include a discussion on related software
    implementations available in all languages (e.g., software
    implementations related to l1 penalized linear regression for instance,
    or even slope problem estimation), highlighting the specific
    contribution of slope.
  \end{quote}

  \begin{quote}
    None of the [package files] can be uncompressed (or used) using gunzip.
  \end{quote}

  Thanks for pointing this out! We have packaged the package files
  correctly now and verified that they can be decompressed without issues.

  \begin{quote}
    It means that, at least a summary method is missing.
  \end{quote}

  Thanks for the suggestion! We have added \verb|summary()| methods for
  both \verb|SLOPE| and \verb|TrainedSLOPE| in the R package.

  \begin{quote}
    Other methods might
    be missing that would avoid exposing the internal structure of objects
    in the replication material [...]
  \end{quote}

  We are not sure we agree that the fields exposed in the objects should be
  considered internal structure (private) nor that it's idiomatic in R
  to hide them behind accessor methods, so we have respectfully chosen not to
  implement such methods. In the case of the SLOPE patterns, this object is
  computed as part of the fitting procedure and can therefore not be (easily)
  externalized to a separate function.

  \begin{quote}
    Using \texttt{example("SLOPE");
      plot(fit)}
    results in an empty plot. It is probably not expected: Can the authors
    look into this?
  \end{quote}

  Thanks! This was indeed a bug. In the new version of the package, we have
  added an entirely different (Cleveland dot plot) visualization of the
  solution for single fits, which should resolve this bug.

  \begin{quote}
    Replication material should be cleaned and simplified a bit, in
    particular to remove hidden configuration files coming from github. Make
    sure to upload only files required to reproduce results of the
    manuscript and to adapt README files to avoid having to use the github
    repository in the instructions (you can of course mention that these
    files are duplicated in a github repository but explain how to run the
    replication code from the local directory).
  \end{quote}

  \begin{quote}
    Following the instructions to run the benchmark given in the README
    file of \verb|benchmark_slope| (after activating a virtual environment for
    Python), we obtained: [...]

    % benchopt install ./benchmark\_slope Traceback (most recent call
    % last): File "myvenv/bin/benchopt", line 8, in
    % \textless module\textgreater sys.exit(benchopt()) \^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{} File
    % "myvenv/lib/python3.12/site-packages/click/core.py", line 1485, in
    % \_\_call\_\_ return self.main(*args,
    % **kwargs) \^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{} File
    % "myvenv/lib/python3.12/site-packages/click/core.py", line 1406, in
    % main rv =
    % self.invoke(ctx) \^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{} File
    % "myvenv/lib/python3.12/site-packages/click/core.py", line 1873, in
    % invoke return
    % \_process\_result(sub\_ctx.command.invoke(sub\_ctx)) \^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{} File
    % "myvenv/lib/python3.12/site-packages/click/core.py", line 1269, in
    % invoke return ctx.invoke(self.callback,
    % **ctx.params) \^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{} File
    % "myvenv/lib/python3.12/site-packages/click/core.py", line 824, in
    % invoke return callback(*args,
    % **kwargs) \^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{}\^{} File
    % "myvenv/lib/python3.12/site-packages/benchopt/cli/main.py", line 452, in
    % install raise RuntimeError( RuntimeError: No conda environment is
    % activated. You should be in a conda environment to use
    % \textquotesingle benchopt install\textquotesingle.\\
  \end{quote}

  \begin{quote}
    If conda has to be used, this should be mentioned in the installation
    instructions. The README file of the replication material indicates that
    this is the case: please make the two files consistent or remove the
    README files of subdirectories.
  \end{quote}

  \begin{quote}
    Also provide a more detailed information
    about setting the conda environment for user not familiar with it.
  \end{quote}

  \begin{quote}
    Running benchopt run in \verb|benchmark_slope| results in errors due to
    missing dependencies (skglm for instance). It would be preferable to
    provide a \verb|requirements.txt| file allowing users to install all required
    dependencies without the need to search what they are.
  \end{quote}

  \begin{quote}
    Trying to run the same benchmark, we also had: [...]
    Can you instruct how this should be solved?
  \end{quote}

  % Traceback (most recent call last): File
  % "replication-material/benchmark\_slope/solvers/slope\_path.py", line 5,
  % in \textless module\textgreater from modules import
  % path\_solver ImportError: cannot import name
  % \textquotesingle path\_solver\textquotesingle{} from
  % \textquotesingle modules\textquotesingle{}
  % anaconda3/lib/python3.12/site-packages/modules.py) \textbar-\/-SlopePath:
  % not installed \textbar-\/-Newt-ALM{[}inner\_solver=standard{]}:
  % done Traceback (most recent call last): File
  % "replication-material/benchmark\_slope/solvers/pgd\_safe\_screening.py",
  % line 6, in \textless module\textgreater from
  % slopescreening.solver.parameters import
  % SlopeParameters ModuleNotFoundError: No module named
  % \textquotesingle slopescreening\textquotesingle \textbar-\/-PGD\_safe\_screening:
  % not installed\\
  % Can you instruct how this should be solved?\\

  \begin{quote}
    Could you please also provide scripts with minimal command lines to
    run the equivalent of \verb|example.R| and/or \verb|real-data.R| with Python and
    Julia?
  \end{quote}

  \begin{quote}
    \verb|help(package = "SLOPE")| shows that not man page titles are not in
    title styles.
  \end{quote}

  We have converted all manual pages to use title case in their titles.

  \begin{quote}
    Probably, \verb|plot.TrainedSLOPE()| should be removed from "Other
    model-tuning".
  \end{quote}

  Thanks! We have removed the duplicate reference.

  \begin{quote}
    It seems that the output of \verb|cvSLOPE| only gives a way to obtain the
    optimal hyperparameters. In R programming, it is standard to also return
    the best model fit with these parameters (see e.g., \verb|?e1071::tune|). Can
    the authors think of an handy way for the user to either obtain this
    trained model from \verb|cvSLOPE| or use the output of \verb|cvSLOPE| in a method that
    would directly train the optimal model?
  \end{quote}

  Thanks for the suggestion! You're right that this was an omission. We have now updated
  \verb|cvSLOPE()| to (optionally and by default), return the fitted model on the full
  data set. We've also introduced a new function (and generic) \verb|refit()| that
  can be used to fit SLOPE onto new data using the parameters selected by
  \verb|cvSLOPE()|.

  \closing{Yours sincerely,\\[1ex]
    Johan Larsson, Ma\l{}gorzata Bogdan, Krystyna Grzesiak, Mathurin Massias, and Jonas Wallin
  }

  \encl{manuscript, software, replication materials, original cover letter}
\end{letter}

\end{document}

\usepackage{libertine}

\title{Cover Letter for Submission to JSS}

\begin{document}

Blabla

\end{document}

