\PassOptionsToPackage{unicode,pdfusetitle}{hyperref}
\PassOptionsToPackage{hyphens}{xurl}
\PassOptionsToPackage{usenames,dvipsnames}{xcolor}

\documentclass[english,a4paper,14pt]{letter}

\usepackage{babel}

% Font settings
\usepackage[T1]{fontenc}
\usepackage{libertine}
\usepackage{inconsolata}

\usepackage{textcomp}
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath}

\usepackage{xcolor}


% tables and lists
\usepackage{enumitem}
\usepackage{booktabs}

\usepackage{graphicx}

\usepackage{xurl}

\usepackage{fancyhdr}
\fancyhead{}
\fancyhead[L]{Larsson, Bogdan, Grzesiak, Massias, Wallin}
\fancyhead[R]{Cover Letter}

\pagestyle{fancy}
\thispagestyle{plain}

\usepackage{hyperref}
\hypersetup{
  pdftitle   = {Cover Letter},
  pdfauthor  = {Johan Larsson, Malgorzata Bogdan, Krystyna Grzesiak, Mathurin Massias, Jonas Wallin},
  colorlinks = true,
  linkcolor  = MidnightBlue,
  citecolor  = MidnightBlue,
  urlcolor   = MidnightBlue,
  breaklinks = true
}

\begin{document}

\begin{letter}{Journal of Statistical Software}
  \opening{Dear editor,}

  We are pleased to submit our manuscript titled ``Efficient Solvers for SLOPE
  in R, Python, Julia, and C++'' for consideration for publication in the
  Journal of Statistical Software.

  This is the second resubmission of a manuscript that was originally submitted
  as number 6061 in November 2025 and then as number 6222. We are thankful for
  the feedback on our previous submission and, yet again, apologize for the
  issues encountered in the submission. At the end of this letter, we provide a
  point-by-point response to the comments raised and outline the changes made to
  address them.

  This paper presents a comprehensive suite of software packages implementing
  efficient solvers for Sorted L-One Penalized Estimation (SLOPE) across four
  programming languages: R, Python, Julia, and C++. The packages support
  generalized linear models with various loss functions and handle diverse data
  structures including dense, sparse, and out-of-memory matrices. Our
  implementation features a highly efficient hybrid coordinate descent
  algorithm and we provide comprehensive benchmarks to showcase its performance
  as well as instructive examples on how to use our software in practice.

  We believe this work makes a significant contribution to the statistical
  software community by providing high-quality, well-documented, and performant
  implementations of SLOPE that are readily accessible across multiple
  programming environments. The packages are already available through standard
  package repositories (CRAN, PyPI, and the Julia General Registry) and have
  been designed with reproducibility and ease of use in mind.

  The submission includes the complete manuscript, all source code for the
  packages, and replication materials that allow readers to
  reproduce all figures and results presented in the paper.

  We confirm that this manuscript has not been published elsewhere and is not
  under consideration by any other journal. All authors have approved the
  manuscript and agree with its submission to JSS.

  Thank you for considering our manuscript. We look forward to your feedback
  and hope that you will find our revised submission suitable for publication in
  the Journal of Statistical Software.

  Here is a point-by-point list the issues raised in the previous submission
  and our answers:

  \begin{quote}
    Two editors tried to run the replication material following the
    installation instruction and we encountered problems: First, an error was found
    trying to setup benchopt with conda environment (the conda environment
    "benchopt" is activated):
  \end{quote}

  This appears to have been a Windows-only issue with R dependencies and Conda
  environment resolution. We have tightened the installation instructions to only
  use the solvers specified in the benchmark configuration and have also updated
  the instructions with clearer guidance on what to do in case of dependency
  issues. One of the authors has verified that the current instructions work
  correctly on a Windows system and we have set up scheduled continuous
  integration tests for all platforms (Windows, Linux, and macOS) to ensure that
  the installation process remains stable. We apologize for the inconvenience
  caused by this issue.

  \begin{quote}
    Thank you for updating the discussion on state-of-the-art methods with other
    SLOPE implementations. There was no misunderstanding from our side in this
    request and we think that it could also have been helpful to cite the most
    important packages implementing other l1 penalization (like the Lasso), with a
    comparison of what SLOPE contains and does not contain in comparison these
    packages.
  \end{quote}

  \begin{quote}
    In the Python package, the class \verb|sortedl1.results.PathResults|
    clearly misses methods to help the user explore the objects. \verb|__str__| is
    not easy to exploit in its current form because the output is too long and
    there is not summary method for instance. Usually, for this type of model
    classes, we also recommend compatibility with scikit-learn so that the
    \verb|model_selection| API can be used directly
    \url{https://scikit-learn.org/stable/api/sklearn.model_selection.html#module-sklearn.model_selection}
  \end{quote}

  Thank you for this helpful suggestion. We
  agree that the previous \verb|PathResults| representation was not
  user-friendly, and have made the following improvements to enhance the exploration of results:

  \begin{itemize}
    \item We added a compact \verb|summary()| method to \verb|sortedl1.results.PathResults|
          to provide key structural information (e.g., number of alphas,
          features, targets,
          and path-related diagnostics) without printing full arrays.
    \item We implemented concise \verb|__str__|/\verb|__repr__| methods
          for \verb|PathResults|,
          so object printing now gives a short, human-readable overview
          instead of a long dump.
    \item For consistency, we also added the same exploration utilities
          to \verb|CvResults|.
  \end{itemize}

  Regarding scikit-learn compatibility: \verb|Slope| is the estimator intended
  to be used with \verb|sklearn.model_selection|, and we added documentation
  showing direct usage with \verb|GridSearchCV|.

  \begin{quote}
    The authors have chosen to implement a refit method for their class
    "trainedSlope". However, this choice does not seem completely valid: it
    requires that the user re-specify x and y during the refit, which can lead to
    inconsistencies with respect to what has been used initially to perform cross
    validation.
  \end{quote}

  Thank you for pointing this out. We agree that requiring users to re-supply
  \texttt{x} and \texttt{y} in \verb|refit()| is problematic, althogh
  we would contend that the goal of refitting typically is to fit a model
  on new data. In any case, we have updated the implementation so that
  \verb|refit.TrainedSLOPE()| now uses the training data stored in the
  \verb|TrainedSLOPE| object by default. Users may still provide new
  \texttt{x}/\texttt{y} explicitly, but both must be given together.

  \begin{quote}
    A similar issue arises with the Python package that does not return the best trained model.
  \end{quote}

  We have updated the API so that
  \verb|Slope.cv(..., refit=True)| returns both cross-validation results and the
  best fitted model, i.e., \verb|(cv_results, best_model)|. The returned
  \verb|best_model| is refit on the same \verb|X, y| passed to the CV call using
  the selected best hyperparameters.

  \closing{Yours sincerely,\\[1ex]
    Johan Larsson, Ma\l{}gorzata Bogdan, Krystyna Grzesiak, Mathurin Massias, and Jonas Wallin
  }

  \encl{manuscript, software, replication materials}
\end{letter}

\end{document}
